Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-22 PAPER.tex
made syntax/typo changes from full-read-through up to Section 2.2

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the $21$ cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher $21$ cm power spectrum upper limit from PAPER-64. This revised result, presented in a companion paper by Kolopanis et al., \textit{submitted}, mostly stems from an 
improved signal loss calculation for loss associated with empirically estimated covariances and supersedes results from 
previously published PAPER analyses.
\end{abstract}

@@

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This tomography technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. Tracing the 21\,cm line as a function of redshift offers a window into the 
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, errors, and bias) that are essential to investigate 
for a robust $21$ cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. More specifically, we develop new techniques for computing signal loss and under-go a rigorous re-analysis of PAPER-64 data in order to present a new power spectrum limit. A companion paper, Kolopanis et al., \textit{submitted}, builds off of the methods in this paper to present PAPER-64 results for multi-redshifts and multiple baselines.

Finally, this paper adds to the growing foundations of lessons which have been documented in  \citet{Patil2016}, \citet{Jacobs2016}, and \cite{Paciga2013} by the LOFAR, MWA and GMRT projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

This paper is organized into two main sections. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al., \textit{submitted}). We conclude in Section \ref{sec:Con}.

@@

There are many choices a $21$ cm data analyst must consider, though the space is significantly constrained by 
the type of instrument being used.  The PAPER instrument in its grid configuration focuses most sensitivity on 
only a few baselines but imaging performance is poor. In this space, signal processing tricks which filter or 
down-weight foreground modes without reference to a sky model have the advantage of being agnostic about 
the sky. However, even within this problem sub-space, questions remain.

 How can time-ordered measurements be combined? How can the 
variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not 
destroying an EoR signal? How can a statistically significant detection of a signal be properly identified? Many common techniques, such as 
averaging data, weighting, bootstrapping, and jackknife testing, address these issues but harbor additional trade-offs. For 
example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing 
some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Though there are many data analysis choices, measuring the statistical $21$ cm power spectrum ultimately requires robust 
methods for determining accurate confidence intervals and rigorous techniques to identify and control systematics.  In this 
paper, we focus on three $21$ cm power spectrum themes that encapsulate this goal and discuss four techniques that interplay 
with each other and impact the themes. We will give brief definitions now, and build intuition for each theme in the sections to 
follow.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a $21$ cm power 
spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important checklist for 
a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Signal loss refers to attenuation of the target cosmological signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if the amount of loss is not quantified accurately, it could lead to false non-detections and overly aggressive upper limits. Computing signal loss correctly has subtle challenges but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on a $21$ cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving a bias is an EoR detection may be the most difficult challenge for $21$ cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must 
pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general $21$ cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter averages visibilities in time to produce a smaller number of more sensitive 
independent samples. However, by adjusting the shape of the filter to match the rate of sidereal motion it can also affect the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by PAPER is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). Using such a technique enables the down-weighting of contaminant modes that obey a different covariance structure from that of cosmological modes. However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two has large impacts on signal loss. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011b}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use a toy data model to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be performed for robust $21$ 
cm results.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or by deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the \citetalias{ali_et_al2015} analysis.

@@

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities for a single baseline in Jy. It has length $N_{t} \times N_{f}$, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along $2$ spatial dimensions (using the flat-sky approximation), and since we are interested in $3$-dimensional Fourier-modes 
we only need to take one Fourier transform of our visibilities along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier transforms them into 
power spectrum space, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\partial\textbf{C}}{\partial p_{\alpha}}$ (which has dimensions ($N_{f}$,$N_{f}$)), or the derivative of the covariance $\textbf{C}$ with respect to the true bandpower $p_\alpha$, where $\alpha$ indexes a waveband in $k_{\parallel}$ ($k_{\parallel}$, a cosmological wavenumber, is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}). $\textbf{R}$ is a weighting matrix with dimensions ($N_{f}$,$N_{f}$) --- as 
an example, inverse covariance weighting would set $\textbf{R} = \textbf{C}^{-1}$ and an uniform-weighted case would use $
\textbf{R} = \textbf{I}$, the identity matrix.

@@

\noindent where $\widehat{\textbf{p}}$ is the estimate of the true power spectrum. The data analyst has a choice for $
\textbf{M}$ (subject to the constraint that window functions integrate to unity) --- for simplicity in this section we choose $\textbf{M}$ to be diagonal ($\textbf{M} \propto \textbf{I}$), although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimate $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss. We will now build intuition into how estimated inverse covariance weighting can give rise to signal loss through the use of a toy model.

We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the previous analysis and our new analysis) we use $\sim$$8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\textbf{p}$(k) (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes throughout Section \ref{sec:SiglossOverview}. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock EoR signal as a random Gaussian signal. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for estimated inverse covariance weighting is then $
\textbf{R} = \widehat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of our toy model dataset $\textbf{x}$ using OQE formalism and $\textbf{R} = \widehat{\textbf{C}}^{-1}$. 
The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the uniform-weighted($\textbf{R} = \textbf{I}$) power spectrum of $
\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$ mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. For an empirically estimated $\widehat{\textbf{C}}$, its eigenvalues differ from those of the true covariance, where the weakest (EoR-dominated) eigenmodes are the most strongly coupled to the data and can lead to the most loss.
Hence, there is negligible signal loss when assigning identical weights to all eigenmodes except the first, since we are not using the relative weights of the weaker eigenmodes.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR dominated modes) to the data. We will next show how this effect is exaggerated by reducing the total number of independent 
samples in a dataset.

>>>>

\begin{abstract}
The Epoch of Reionization (EoR) is an uncharted era in our Universe's history during which the first stars and 
galaxies led to the ionization of neutral hydrogen in the intergalactic medium. There are many experiments investigating the 
EoR by tracing the 21\,cm line of neutral hydrogen, a signal which is very faint and difficult to isolate. With a new generation 
of instruments and a statistical power spectrum detection in our foreseeable future, it has become increasingly important to 
develop techniques that help maximize sensitivity while validating results. Additionally, it is imperative to understand the trade-offs 
between different methods and their effects on power spectrum estimates. In this paper, we focus on three major 
themes --- signal loss, power spectrum error estimation, and bias in measurements. We describe techniques that affect 
these themes using both toy models and data taken by the 64-element configuration of the Donald C. Backer Precision Array 
for Probing the Epoch of Reionization (PAPER). In particular, we highlight how detailed investigations of these themes have 
led to a revised, higher 21\,cm power spectrum upper limit from PAPER-64. This revised result, presented in a companion paper by Kolopanis et al. (\textit{submitted}), mostly stems from an 
improved signal loss calculation for loss associated with empirically estimated covariances and supersedes results from 
previously published PAPER analyses.
\end{abstract}

@@

One promising technique to probe the EoR is to target the 21\,cm wavelength signal that is emitted and absorbed by neutral hydrogen via 
its spin-flip transition (\citealt{furlanetto_et_al2006}; \citealt{barkana_and_loeb2008}; \citealt{morales_and_wyithe2010}; \citealt{pritchard_and_loeb2010}; \citealt{pritchard_loeb2012}). This technique is powerful because it can be observed both spatially and as a function of redshift --- that is, the wavelength 
of the signal reaching our telescopes can be directly mapped to a distance from where the emission originated before 
stretching out as it traveled through expanding space. 21\,cm tomography offers a unique window into the 
evolution of ionization, temperature, and density fluctuations.

@@

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, error estimation, and bias) that are essential to investigate 
for a robust 21\,cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. A companion paper, Kolopanis et al. (\textit{submitted}), builds off of the methods in this paper to present PAPER-64 results for multiple redshifts and baseline types.

Finally, this paper adds to the growing foundations of lessons which have been documented in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

This paper is organized into two main halves. In Section \ref{sec:Themes} we introduce the three themes of our focus, using a 
toy model to develop intuition for each one. In Section \ref{sec:CaseStudy} we present a case study into each theme using data 
from the PAPER-64 array, highlighting key changes from the previously published result in \citet{ali_et_al2015}, henceforth known as \citetalias{ali_et_al2015}, which have led to a 
revised PAPER-64 power spectrum result (Kolopanis et al. (\textit{submitted})). We conclude in Section \ref{sec:Con}.

@@

There are many choices a 21\,cm data analyst must consider. %For PAPER, its grid configuration allows the use of many signal processing tricks to down-weight foregrounds without reference to a sky model. However, even with the advantage of being agnostic about the sky, many questions still remain.
%, though the space is significantly constrained by 
%the type of instrument being used.  The PAPER instrument in its grid configuration focuses most sensitivity on 
%only a few baselines but imaging performance is poor. In this space, signal processing tricks which filter or 
%down-weight foreground modes without reference to a sky model have the advantage of being agnostic about 
%the sky. However, even within this problem sub-space, questions remain.
How can time-ordered measurements be combined? How can the 
variance of the data be estimated? In what way(s) can the data be weighted to suppress contaminated modes while not 
destroying an EoR signal? How can a statistically significant detection of a signal be properly identified? Many common techniques, such as 
averaging data, weighting, bootstrapping, and jackknife testing, address these issues but harbor additional trade-offs. For 
example, an aggressive filtering method may succeed in eliminating interfering systematics but comes at the cost of losing 
some EoR signal. A chosen weighting scheme may maximize sensitivity but fail to suppress foregrounds.

Though there are many data analysis choices, measuring the statistical 21\,cm power spectrum ultimately requires robust 
methods for determining accurate confidence intervals and rigorous techniques to identify and control systematics.  In this 
paper, we focus on three 21\,cm power spectrum themes that encapsulate this goal and discuss four techniques that interplay 
with each other and impact the themes. We will give brief definitions now, and build intuition for each theme in the sections to 
follow.

@@

A deep understanding of the following three themes is essential for the accuracy and interpretation of a 21\,cm power 
spectrum result. Stemming from a re-analysis of PAPER-64 data, we believe these themes serve as an important checklist for 
a rigorous power spectrum analysis.
\begin{itemize}
\item \textbf{Signal Loss} (Section \ref{sec:SiglossOverview}): Signal loss refers to attenuation of the target cosmological signal 
in a power spectrum estimate. Certain analysis techniques can cause this loss, and if the amount of loss is not quantified accurately, it could lead to false non-detections and overly aggressive upper limits. Computing signal loss correctly has subtle challenges but is necessary to ensure the accuracy of any result. 
\item \textbf{Error Estimation} (Section \ref{sec:ErrorOverview}): Confidence intervals on a 21\,cm power spectrum result 
determine the difference between a detection and a null result, which have two very different implications. Errors can be 
estimated in a variety of ways, and we will discuss a few of them.
\item \textbf{Bias} (Section \ref{sec:BiasOverview}): There are several possible sources of power offset in a visibility 
measurement that can show up as a detection in a power spectrum, such as bias from noise and foregrounds. In particular, a 
successful EoR detection would also imitate a bias. Proving a bias is an EoR detection may be the most difficult challenge for 21\,cm 
analyses, as it is crucial to be able to distinguish a detection of foreground leakage, for example, from that of EoR. In this paper 
we will highlight some sources of bias, discuss ways to mitigate their effects, and describe tests that a true EoR detection must 
pass.
\end{itemize}

@@

The following techniques each have advantages when it comes to maximizing sensitivity and understanding systematics in 
data. However, some have limitations, and we will discuss circumstances in which there are trade-offs. We choose to focus on 
these four techniques because they represent major steps in PAPER's power spectrum pipeline, with several of them also 
being standard steps in general 21\,cm analyses.
\begin{itemize}
\item \textbf{Fringe-rate filtering:} Fringe-rate filtering is an averaging scheme for time-ordered data 
(\citealt{parsons_et_al2016}). Broadly, a fringe-rate filter averages visibilities in time to produce a smaller number of more sensitive 
independent samples. However, when adjusting the shape of the filter to match the rate of sidereal motion, it can also affect the presence 
of foregrounds and systematics. We explain the trade-offs of filtering in more detail in Section \ref{sec:toymodel_frf}.
\item \textbf{Weighting:} A dataset can be weighted to emphasize certain features and minimize others. One particular flavor of 
weighting employed by previous PAPER analyses is inverse covariance weighting in frequency, which is a generalized version of inverse variance 
weighting that also takes into account frequency correlations (\citealt{liu_tegmark2011}; \citealt{dillon_et_al2013a}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2014}; \citealt{dillon_et_al2015}). Using such a technique enables the down-weighting of contaminant modes that obey a different covariance structure from that of cosmological modes. However, a challenge of inverse covariance 
weighting is in estimating a covariance matrix that is closest to the true covariance of the data; the discrepancy between the two has large impacts on signal loss. We investigate the impact of different types of weighting on signal loss in Section \ref{sec:SiglossOverview}.
\item \textbf{Bootstrapping:} In addition to using theoretical models for covariance matrices and theoretical error estimation 
methods, bootstrapping is one way to estimate errors. Namely, bootstrapping is a useful method for estimating errors of a dataset from 
itself (\citealt{andrae2010}). By randomly drawing many subsamples of the data, we obtain a sense of its inherent variance, though there are subtleties to 
consider such as the independence of values in a dataset. We explore this potential pitfall of bootstrapping in Section \ref{sec:ErrorOverview}.
\item \textbf{Jackknife testing:} A resampling technique useful for estimating bias, jackknives can be taken along different 
dimensions of a dataset to cross-validate results. In particular, null tests can be used to verify whether results are free of 
systematics, as done with CO power spectra (\citealt{keating_et_al2016}) and CMB measurements (see e.g. \citealt{ade_et_al2008}; \citealt{chiang_et_al2010}; \citealt{bischoff_et_al2011}; \citealt{das_et_al2011b}; \citealt{araujo_et_al2012}; \citealt{crites_et_al2015}; \citealt{ade_et_al2016}; \citealt{ade_et_al2017}; \citealt{sherwin_et_al2017}). An EoR detection must pass both jackknife and null tests, which we highlight in Section \ref{sec:JackknifeOverview}.
\end{itemize}

In the next three subsections, we study each theme in depth, focusing on how power spectrum technique trade-offs affect each. 
We use a toy data model to develop intuition into why certain analysis choices may be appealing and discuss ways in which 
they are limited. We highlight problems that can arise regarding each theme and offer suggestions to mitigate the issues. 
Ultimately, we show that rigorous investigations into signal loss, error estimation, and bias must be performed for robust 21\,cm results.

@@

Signal loss can arise in a variety of ways in the analysis pipeline, such as by fitting a polynomial during 
spectral calibration, applying a delay-domain filter, or deriving weights from data and applying them to itself. Here we focus on signal loss associated with 
applying a weighting matrix to data, a loss that can be significant depending on the choice of weighting and one that was 
previously underestimated in the \citetalias{ali_et_al2015} analysis.

@@

We begin with our data vector, $\textbf{x}$, which contains our measured visibilities for a single baseline in Jy. It has length $N_{t} \times N_{f}$, but in practice we manipulate it as an array with dimensions ($N_{f}, N_{t}$), where 
$N_{t}$ is the number of time integrations and $N_{f}$ is the number of frequency channels. Visibilities are measurements of 
the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier-modes 
we only need to take one Fourier transform of our visibilities along the line-of-sight dimension. We do this when forming the un-normalized power spectrum estimate $\widehat{q}_{\alpha}$:

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier transforms them into 
power spectrum space, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is formally evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\partial\textbf{C}}{\partial p_{\alpha}}$ (which has dimensions ($N_{f}$,$N_{f}$)), or the derivative of the covariance $\textbf{C}$ with respect to the true bandpower $p_\alpha$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the 
Fourier-dual to frequency under the delay approximation (\citealt{parsons_et_al2012b})). $\textbf{R}$ is a weighting matrix with dimensions ($N_{f}$,$N_{f}$) --- as 
an example, inverse covariance weighting would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $
\textbf{R} \equiv \textbf{I}$, the identity matrix.

@@

\noindent where $\widehat{\textbf{p}}$ is the estimate of the true power spectrum $\textbf{p}$. The data analyst has a choice for $
\textbf{M}$ (subject to the constraint that window functions integrate to unity) --- for simplicity in this section we choose $\textbf{M}$ to be diagonal ($\textbf{M} \propto \textbf{I}$), although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimates $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss. We will now build intuition into how empirically estimated inverse covariance weighting can give rise to signal loss through the use of a toy model.

We construct a simple dataset that contains 2-dimensional data (representing visibility data with $100$ time 
integrations and $20$ frequency channels). This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim$$8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down for this toy model with no loss of 
generality.

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift-scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $\textbf{p}(k)$ (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$, to which we apply different weighting schemes throughout Section \ref{sec:SiglossOverview}. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss1.png}
	\caption{Our toy model dataset to which we apply different weighting schemes to in order to investigate signal loss. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock EoR signal as a random Gaussian signal. We add the two together to form $\textbf{x} = 
\textbf{x}_{FG} + \textbf{x}_{EoR}$. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

\noindent assuming $\langle\textbf{x}\rangle_{t} = 0$ (a reasonable assumption since fringes average to $0$ over a sufficient 
amount of time), where $\langle \rangle_{t}$ denotes a finite average over time. The weighting matrix for our empirically estimated inverse covariance weighting is then $
\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$.

First, we compute the power spectrum of our toy model dataset $\textbf{x}$ using OQE formalism and $\textbf{R} \equiv \widehat{\textbf{C}}^{-1}$. 
The result is shown in green in the left plot of Figure \ref{fig:toy_sigloss3}. Also plotted in the figure are the uniform-weighted ($\textbf{R} \equiv \textbf{I}$) power spectrum of the individual components $
\textbf{x}_{FG}$ (blue) and $\textbf{x}_{EoR}$ (red).

\begin{figure*}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss3.png}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,height=0.33\textwidth]{plots/toy_sigloss4.png}
	\caption{Resulting power spectrum estimates for the toy model simulation described in Section \ref{sec:toymodel} --- 
foregrounds only (blue), EoR only (red), and the weighted FG + EoR dataset (green). The power spectrum of foregrounds peaks at a $k$-mode based on the frequency of the sinusoid used to create the mock FG signal. In the two panels, we compare empirically estimated inverse covariance weighting 
where $\textbf{C}$ is derived from the data (left), and projecting out the first eigenmode only (right). In the former 
case, signal loss arises from the coupling of the eigenmodes of $\widehat{\textbf{C}}$ to the data. For an empirically estimated $\widehat{\textbf{C}}$, its eigenvalues differ from those of the true covariance, where the weakest (EoR-dominated) eigenmodes are the most strongly coupled to the data and can lead to the most loss.
Hence, there is negligible signal loss when assigning identical weights to all eigenmodes except the first, since we are not using the relative weights of the weaker eigenmodes.}
	\label{fig:toy_sigloss3}
\end{figure*}

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR-dominated modes) to the data. We will next show how this effect is exaggerated by reducing the total number of independent 
samples in a dataset.
